---
title: Tidy Tuesdays - 2020-04-07
subject: Tour de France
author: Jonny White
website: jonnywhitedatascience.github.io
date: 13/04/2020
---

Load libraries
```{r}

library(tidyverse)
library(extrafont)
library(remotes)
library(tdf)

loadfonts(device="win", quiet = TRUE)

#load my theme from external file
source("https://raw.githubusercontent.com/jonnywhitedatascience/TidyTuesday/master/theme/jwtheme.R")
```


The below code first of all creates a simple summary tibble *national_data* which includes nationality and a count.  Next it plots on column chart using ggplot and geom_col, with data labels from geom_text.  It then assigns various plot labels using labs, rotates the plot by 90 degrees and ends with my chatrt theme (which is a simple version of those found in the Economist)

```{r}

nationality_data <- editions %>% 
  group_by(nationality) %>%
  summarize(count=n())
  
nationality_data %>%
  ggplot( 
    aes(
      x=reorder(nationality, count),
      y=count)
    )+
  geom_col()+
  geom_text(
    aes(
      label=count),
      nudge_y=-0.5,
      size=3,
      colour="white")+
  labs(
    title="Tour de France",
    subtitle="edition winner by nationality",
    x="nationality",
    y="edition wins",
    caption = "source:  @rushworth_a"
  )+
  coord_flip()+
  jwtheme
```

Next i want to create a matrix view of edition winners by nationality and year to see if there have been particular times of success for certain countries.  From the little i know of the Tour, Lance Armstrong was particularly successful, so this plot should demonstrate American dominance in the late 90's / early 00's?

```{r}
#prepare the data
year_nationality_data<-
  transmute(
    editions,
    year=as.numeric(format(as.Date(start_date, format-"%Y-%m-%d"),"%Y")),
    nationality,
    winner_name,
    win_count=1,
  ) %>%
  group_by(nationality)%>%
  mutate(latest_win_year=max(year), total_wins=sum(win_count))%>%
  ungroup

#plot the chart

plot2<-
  ggplot(
  data=year_nationality_data,
  mapping=aes(
    x=reorder(nationality, latest_win_year),
    y=year,
    label=winner_name)
  )+
  
  geom_tile(
    mapping=aes(
      fill=win_count
    ),
    colour="light blue",
  )+
  
  labs(
    title="Tour de France",
    subtitle="Nationality of winner by year, sorted by most recent winning year",
    x="nationality",
    y="",
    caption = 
    "Author: Jon White \ @_jondwhite \n Tidy Tuesday 07/04/2020 | Source:  Alastair Rushworth \ @rushworth_a"
  )+
    
  scale_y_continuous(
    breaks = seq(1900,2020, by=10), 
    )+
    
  coord_flip()+
    
  jwtheme+
    
  theme(
    panel.grid.minor.x= element_blank()
  )

plot2
```

